---
title: "Excercise 4 - Centrality & Efficiency"
output: pdf_document
author: Hugo Garcia (260791363)
date: March 4th, 2023
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(arrow)
library(tidygraph)
library(ggraph)
```

## Load data
Load the following data:
  + applications from `app_data_sample.parquet`
  + edges from `edges_sample.csv`

```{r}
# change to your own path!
data_path <- "C:/Users/hugog/Desktop/Exercise 4/"
applications <- read_parquet(paste0(data_path,"app_data_sample.parquet"))
edges <- read_csv(paste0(data_path,"edges_sample.csv"))

applications
edges
```


## Get gender for examiners
``` {r}
library(gender)
examiner_names <- applications %>% 
  distinct(examiner_name_first)
examiner_names
```

## Get a table of names and gender
```{r}
examiner_names_gender <- examiner_names %>% 
  do(results = gender(.$examiner_name_first, method = "ssa")) %>% 
  unnest(cols = c(results), keep_empty = TRUE) %>% 
  select(
    examiner_name_first = name,
    gender,
    proportion_female
  )
examiner_names_gender
```


```{r}
# remove extra colums from the gender table
examiner_names_gender <- examiner_names_gender %>% 
  select(examiner_name_first, gender)
# joining gender back to the dataset
applications <- applications %>% 
  left_join(examiner_names_gender, by = "examiner_name_first")
# cleaning up
rm(examiner_names)
rm(examiner_names_gender)
gc()
```


```{r}
library(wru)
examiner_surnames <- applications %>% 
  select(surname = examiner_name_last) %>% 
  distinct()
examiner_surnames
```

We'll follow the instructions for the package outlined here [https://github.com/kosukeimai/wru](https://github.com/kosukeimai/wru).

NOTE: I was getting errors running the original code block for examiner_race. I tried updating packages, and debugging for a long time but I believe it is my computer's software/environment setup preventing me. I asked for the csv output from a peer and am importing it instead.
Original code bloack:
  examiner_race <- predict_race(voter.file = examiner_surnames, surname.only = T) %>% 
    as_tibble()
  examiner_race

```{r race-2}
examiner_race <- read_csv(paste0(data_path,"examiner_race.csv"))
examiner_race
```

```{r}
examiner_race <- examiner_race %>% 
  mutate(max_race_p = pmax(pred.asi, pred.bla, pred.his, pred.oth, pred.whi)) %>% 
  mutate(race = case_when(
    max_race_p == pred.asi ~ "Asian",
    max_race_p == pred.bla ~ "black",
    max_race_p == pred.his ~ "Hispanic",
    max_race_p == pred.oth ~ "other",
    max_race_p == pred.whi ~ "white",
    TRUE ~ NA_character_
  ))
examiner_race
```

Join the data back to the applications table.
```{r}
# removing extra columns
examiner_race <- examiner_race %>% 
  select(surname,race)
applications <- applications %>% 
  left_join(examiner_race, by = c("examiner_name_last" = "surname"))
rm(examiner_race)
rm(examiner_surnames)
gc()
```
## Add Tenure
```{r}
library(lubridate) # to work with dates
examiner_dates <- applications %>% 
  select(examiner_id, filing_date, appl_status_date) 
examiner_dates
```

```{r}
examiner_dates <- examiner_dates %>% 
  mutate(start_date = ymd(filing_date), end_date = as_date(dmy_hms(appl_status_date)))
```

```{r}
examiner_dates <- examiner_dates %>% 
  group_by(examiner_id) %>% 
  summarise(
    earliest_date = min(start_date, na.rm = TRUE), 
    latest_date = max(end_date, na.rm = TRUE),
    tenure_days = interval(earliest_date, latest_date) %/% days(1)
    ) %>% 
  filter(year(latest_date)<2018)
examiner_dates
```
```{r}
applications <- applications %>% 
  left_join(examiner_dates, by = "examiner_id")
rm(examiner_dates)
gc()
```

## Pre-Processing

```{r}
library(tidyverse)
# Select applications have been either abandoned or issued
abandoned_apps = applications[!is.na(applications$abandon_date),]
issued_apps = applications[!is.na(applications$patent_issue_date),]

# Rename and remove unnecessary columns
abandoned_apps = abandoned_apps %>% rename(end_date = abandon_date) %>% select(-c('patent_issue_date'))
issued_apps = issued_apps %>% rename(end_date = patent_issue_date) %>% select(-c('abandon_date'))
issued_apps$issued = 1
abandoned_apps$issued = 0

# Combine abandoned and issued dates
applications = rbind(abandoned_apps, issued_apps)
rm(abandoned_apps, issued_apps)
```

## Calculate Application Processing Times

```{r}
app_proc_time = applications$end_date - applications$filing_date
app_proc_time = as.numeric(app_proc_time)
summary(app_proc_time)
```
```{r}
# There were some errors in the dates which led to negative app_proc_time, remove these
applications$app_proc_time = app_proc_time
applications = applications[applications$app_proc_time >=0, ]
```


## Calculate Centrality Scores for Each Examiner

```{r}
edges = edges %>% rename(to = alter_examiner_id,
                     from = ego_examiner_id)

# This is a directed  network, so Directed=TRUE
graph = as_tbl_graph(x = edges[c('to','from')], directed = TRUE , mode = 'out')
nodes = graph %>%
  activate(nodes) %>%
  mutate(degree = centrality_degree(), 
         closeness = centrality_closeness(),
         betweenness = centrality_betweenness()) %>% 
  rename(examiner_id = name) %>%  data.frame()
applications$examiner_id = as.character(applications$examiner_id)
applications = applications %>% left_join(nodes, by = 'examiner_id')
```
## Drop NA Vaues and Set some Variables as factor

```{r}
attach(applications)
applications = applications %>% drop_na(gender)
applications = applications %>% drop_na(race)
applications = applications %>% drop_na(degree)
applications = applications %>% drop_na(closeness)
applications = applications %>% drop_na(betweenness)
applications = applications %>% drop_na(tenure_days)

applications$gender = as.factor(applications$gender)
applications$race = as.factor(applications$race)

```

## Use a Linear Regression Model to Estimate the Relationship between Centrality and Application Processing Times 
### Controlling for other characteristics of the examiner which might influence that relationship

```{r}
applications_lm <- applications
lm = lm(applications_lm$app_proc_time ~ 
            applications_lm$degree + 
            applications_lm$closeness + 
            applications_lm$betweenness + 
            applications_lm$gender + 
            applications_lm$race + 
            applications_lm$issued +
          applications_lm$tenure_days)
summary(lm)

```

## Results 1

All independent variables are significant to predict the number of days to process a patent, except "Race - Other", and to a lesser degree of significance "Race- Hispanic", which is likely a result of there being relatively fewer Hispanic Examiners and thus fewer applications in total on their behalf.
 
The multiple R-squared value of 0.01011 indicates that the independent variables explain just 1% of the variation in the applications processing times. 
We also see that Closeness Centrality significantly influences application processing time whereby each additional degree of closeness decreases the application processing time by 118 days. Similarly for Degree of Centrality, but to a lesser extent, each additional Degree of Centrality decreases application processing time by about 20 days. 
Lastly, the Tenure of an examiner significantly influences their application processing times whereby each  

In addition, based on the regression model's output we can also make comparisons between examiner characteristics and how they relate to their application processing times:

- Male Examiners take about 35 days longer, on average, to process applications than female examiners
- White Examiners take about 58 days less, on average, to process applications.

## Linear Regression Including Interaction Variables for Gender and Centrality

```{r}
lm_2 = lm(applications_lm$app_proc_time ~ 
            applications_lm$degree + 
            applications_lm$closeness + 
            applications_lm$betweenness + 
            applications_lm$gender + 
            applications_lm$race + 
            applications_lm$tenure_days+
            applications_lm$issued + 
            (applications_lm$gender*applications_lm$degree) + 
            (applications_lm$gender*applications_lm$betweenness) + 
            (applications_lm$gender*applications_lm$closeness))
summary(lm_2)
```

## Results 2

We see similar results than in the regression above for the significance of the independent variables.
Regarding the interactions between gender and centrality measures, all are statistically significant.

